---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# robustGMM

<!-- badges: start -->
<!-- badges: end -->

The goal of robustGMM is to implement the robust Gaussian Mixture Model (GMM) estimation based on [Fujisawa and Eguchi, 2006](https://doi.org/10.1016/j.jspi.2005.03.008).

## Installation

You can install the development version of robustGMM from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ge-li/robustGMM")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r}
## Generate a 2-component mixture 
library(robustGMM)
set.seed(404)
lambda <- c(0.25, 0.75)
mu <- c(0, 4)
sigma <- c(1, 1)
x <- rnormix(n=100, lambda, mu, sigma)
x[which.max(x)] <- 10 # outlier
plot(density(x))
```

The standard EM algorithm will give estimation results as follows:
```{r}
standard_mod <- mixtools::normalmixEM(x, lambda, mu, sigma)
standard_mod$lambda
standard_mod$mu
standard_mod$sigma
```

The robust EM algorithm in this package will give:
```{r}
robust_mod <- robustGMM(x, lambda, mu, sigma, beta=0.1, verbose=TRUE)
robust_mod$lambda
robust_mod$mu
robust_mod$sigma
```

## Selection of tuning parameter 
```{r}
betas <- seq(0.01, 0.2, by=0.01)
div <- numeric(0)
for (beta in betas) {
  div <- c(div, loo_cvm_div(x, lambda, mu, sigma, beta))
}
```

```{r}
plot(betas, div, xlab="beta", ylab="empirical Cramer–von Mises divergence", 
     main="Selection of optimal tuning parameter on contaminated synthetic data.")
```

```{r}
res <- purrr::map_dfr(betas, ~unlist(robustGMM(x, lambda, mu, sigma, .)))
res
```

```{r}
par(mfrow=c(2, 2))
plot(mu1 ~ beta, res, ylim=c(mu[1]-0.2, mu[1]+0.2))
abline(h = mu[1], col="gold")
abline(v = betas[which.min(div)], col="darkgreen", lty=2)
plot(mu2 ~ beta, res, ylim=c(mu[2]-0.2, mu[2]+0.2))
abline(h = mu[2], col="gold")
abline(v = betas[which.min(div)], col="darkgreen", lty=2)
plot(sigma1 ~ beta, res, ylim=c(sigma[1]-0.2, sigma[1]+0.2))
abline(h = sigma[1], col="gold")
abline(v = betas[which.min(div)], col="darkgreen", lty=2)
plot(sigma2 ~ beta, res, ylim=c(sigma[2]-0.2, sigma[2]+0.2))
abline(h = sigma[2], col="gold")
abline(v = betas[which.min(div)], col="darkgreen", lty=2)
```
